---
title: 'ANR BISE: Analyzing trophic webs accross agricultural sites with different
  treatments.'
output:
  html_document:
    df_print: paged
encoding: UTF-8
---

```{r setup }
knitr::opts_knit$set(root.dir = "D:/Knowledge-based/BISE/meta-bio-net/Bionet_V2/")
knitr::opts_knit$set(warning = FALSE, include=FALSE)
```

#Setting global parameters
```{r warning=FALSE, include=FALSE}
workdir="D:/Knowledge-based/BISE/meta-bio-net/Bionet_V2/"
source(paste(workdir,"Init.R",sep=""))
### Files
file_kwBase=paste(workdir,"Knowledge/KB.csv",sep="")
file_occurBase=paste(workdir,"Data/normalized_metabar_out.csv",sep="")
experimental_setting=paste(workdir,"Data/experimental_setting.csv",sep="")

file_raw_adjacency=paste(workdir,"Applications/BISE/out/metaweb_adj/bin_adjacency.csv",sep="")
file_raw_weighted_adjacency=paste(workdir,"Applications/BISE/out/metaweb_adj/weighted_adjacency.csv",sep="")

file_sbm_groups=paste(workdir,"Applications/BISE/out/sbmmetaweb_adj/trophic_groups_memberships.csv",sep="")
file_sbm_adjacency=paste(workdir,"Applications/BISE/out/sbmmetaweb_adj/trophic_groups_adjacency.csv",sep="")

### Data structures
##Knowledge base: what columns to fetch
kb_resource_metadata=c("resource","resource_name","resource_rank","res_surf","res_subsurf","res_soil")
kb_consumer_metadata=c("consumer","consumer_name","consumer_rank","cons_surf","cons_subsurf","cons_soil")
taxa_metadata_colnames=c("key","name","rank","surf","subsurf","soil")

edge_ends=c("resource","consumer")
edge_attributes=c("type","cooccur")

### Level of organization of experimental setting ###
levels=c("Site","Parcelle","Echantillon","Repet")

##Metabar output
###what columns define taxa information
metabar_taxa_metadata=c("key","verbatimScientificName","broad")
###what columns contain occurrence data
sample_ids=function(x){
  return(colnames(x)[grep("BISE_*",colnames(x))])
}

### Thresholds
min_prob=0.1  ###Minimum probability of co-occurrence (habitat sharing) to accept interaction
min_occur=1 ###Minimum occurrence count to consider actual presence
sbm_done=0
beta_done=1
betapt_done=1
sbmbetapt_done=0

### Output files
### Saving ###
file_alphametrics_taxa=paste(workdir,"Applications/BISE/out/taxa_topo_energetics_metrics.csv",sep="")
file_alphametrics_groups=paste(workdir,"Applications/BISE/out/groups_topo_energetics_metrics.csv",sep="")
file_alltaxa_rds=paste(workdir,"Applications/BISE/out/taxa_analysis.rds",sep="")
file_beta=paste(workdir,"Applications/BISE/out/beta/betadiv.csv",sep="")
file_beta_pt=paste(workdir,"Applications/BISE/out/beta/betadiv_pt_",sep="")
file_sbmbeta_pt=paste(workdir,"Applications/BISE/out/beta/sbmbetadiv_pt.csv",sep="")
file_redund_fct=paste(workdir,"Applications/BISE/out/func_redund/redund_fct.csv",sep="")
file_agg_func_occur=paste(workdir,"Applications/BISE/out/func_redund/occur_fct.csv",sep="")
```

#Knowledge base load and preparation
First, we load the knowledge base of ecological interactions.

```{r}
### Knowledge base ###
edge_kb= read.csv2(file_kwBase,dec=".")
cat("Knowledge base contains: ",dim(edge_kb)[1],"edges \n",sep=" ")
```

Some statistics about our knowledge base so far ...

```{r}
### Taxa related knowledge ###
#edge_kb$resource_rank%in%c("FAMILY","GENUS","SPECIES","RESOURCE")
resource_metadata=unique(edge_kb[,kb_resource_metadata])
consumer_metadata=unique(edge_kb[,kb_consumer_metadata])
colnames(resource_metadata)=colnames(consumer_metadata)=taxa_metadata_colnames
taxa_kb=unique(rbind(resource_metadata,consumer_metadata))

## Delete unused data ##
rm(resource_metadata)
rm(consumer_metadata)

cat("Knowledge base covers: ",dim(taxa_kb)[1],"taxa \n",sep=" ")
#summary(taxa_metadata)
```

#Occurrence dataset load and preparation
Load motu occurrence dataset (the file has been preprocessed previously so that motu names match the taxonomy we're using)

```{r}
### MOTU occurrence dataset ###
raw_metabar=read.csv2(file_occurBase)
### Occurrence columns ###
pt_names=sample_ids(raw_metabar)
metabar=subset(raw_metabar,rowSums(raw_metabar[,pt_names])>0)
cat("MOTU dataset contains: ",dim(metabar)[1],"motus \n",sep=" ")
rm(raw_metabar)
fac<-read.csv2(experimental_setting)
```

Some MOTUs are assigned to taxa not covered in our knowledge base, we decided to drop them. However, a vast majority of MOTUs are well covered. 
```{r}
### Filtering MOTUs not covered in knowledge base  ###
list_keys=unique(metabar[,metabar_taxa_metadata])
motu_uncovered=subset(list_keys,!(key %in% edge_kb$resource) & !(key%in% edge_kb$consumer))
motu_covered=subset(list_keys,key %in% edge_kb$resource | key%in% edge_kb$consumer)
cat("MOTU coverage in knowledge base: ",100*dim(motu_covered)[1]/dim(list_keys)[1],"%\n")
rm(list_keys)
rm(motu_uncovered)
taxa_metadata=merge(motu_covered,taxa_kb,by="key")
#taxa_metadata=subset(o,rank%in%c("FAMILY","GENUS","SPECIES","RESOURCE"))
rm(motu_covered)
rm(taxa_kb)
```

We'll subset part of the knowledge base that is of interest to the community composition we have in our hands.
```{r}
sub_edges=subset(edge_kb,edge_kb$resource%in%taxa_metadata$key & edge_kb$consumer%in%taxa_metadata$key)
  cat("Covered taxa in knowledge base are involved in ",dim(sub_edges)[1]," known interactions with proportions of trophic and parasitism of respectively: \n",100*(summary(sub_edges[,c("type")])/dim(sub_edges)[1]))

rm(edge_kb)
```

We'll now proceed to aggregating MOTUs assigned to the same taxon by summing their occurrences (read counts) in each sampling point. 
```{r}
### Aggregating MOTU data to the occurrence point ###
cat("Selecting occurrences of covered taxa in the knowledge base \n")
submetabar=metabar[metabar$key %in%taxa_metadata$key,c("key",pt_names)]
rm(metabar)
agg_occur=aggregate(submetabar[,pt_names],
                    by=list(
                      key=submetabar$key),
                      #verbatimScientificName=submetabar$verbatimScientificName,
                      #broad=submetabar$broad),
                      FUN=sum)
rm(submetabar)
```

Drop some plot statistic of aggregated occurrences here. 

# Metaweb construction workflow
So, at this point we have all it takes to run the automated workflow, that is a nice knowledge base of interactions edge_kb and a ready to use occurrence dataset agg_occur all involving the subset of taxa described in taxa_metadata.  

## Building the metaweb and computing metrics on projected networks
```{r}
alltaxa_analysis=meta_bionet_workflow(sub_edges,agg_occur,taxa_metadata,
                          name="BISE",pt_names=pt_names,
                          rank_filter=c("FAMILY","GENUS","SPECIES","RESOURCE"),
                               edge_ends=edge_ends,edge_attributes=edge_attributes,
                               weight_att=2,type_att=1,min_prob=min_prob,min_occur=min_occur,
                               verb=0)

saveRDS(alltaxa_analysis,file = file_alltaxa_rds)
### Saving statistics ###
write.csv2(cbind(fac,alltaxa_analysis$TopoEnMetrics),file=file_alphametrics_taxa,row.names = F)
```

## Network betadiversity metrics
<!-- Here, we analyze betadiversity between observations, at repetition level -->
<!-- ```{r} -->
<!-- eta=0 -->
<!-- if(beta_done==0){ -->
<!--   betadivdf=betanet(gList=alltaxa_analysis$SubNetworks, -->
<!--                     file_beta = file_beta,pt_names = lnames,div='P',eta=eta) -->
<!--   beta_done=1 -->
<!-- }else{ -->
<!--   betadivdf=read.csv2(file_beta,row.names = 1) -->
<!-- } -->
<!-- ``` -->

Here, we analyze betadiversity between points ...

First, we aggregate occurrences to the point level
```{r}
lnames=unlist(lapply(1:64, function(x) sprintf("BISE_%02d",x)))
lgroups=rep(lnames,each=4)
names(lgroups)=pt_names
agg_occur_pt<-aggregate_occur(agg_occur,pt_names,lgroups)
```

Then, we project the metaweb to the new occurrences:
```{r}
pointlevel_analysis=meta_bionet_workflow(sub_edges,agg_occur_pt$df,taxa_metadata,
                               name="BISE_PT",pt_names=lnames,
                          rank_filter=c("FAMILY","GENUS","SPECIES","RESOURCE"),
                               edge_ends=edge_ends,edge_attributes=edge_attributes,
                               weight_att=2,type_att=1,min_prob=min_prob,min_occur=min_occur,
                               verb=0)
```

Finally, we compute the betadiversity between the 64 point-level local networks
```{r message=FALSE}
eta=c(0,1,2)
divs=c('P','L')
if(betapt_done==0){
for (div in divs){
  for (e in eta){
      betadiv_ptdf=betanet(gList=pointlevel_analysis$SubNetworks,
                           file_beta = paste(file_beta_pt,"_",e,"_",div,".csv",sep=""),
                           pt_names = lnames,div=div,eta=e)
  }
}
  betapt_done=1
}else{
  for(div in divs){
    for(e in eta){
      betadiv_ptdf=read.csv2(paste(file_beta_pt,"_",e,"_",div,".csv",sep=""),row.names = 1)
      x=pcoa_plot(D=as.dist(betadiv_ptdf),Y=NULL,corr="none")
    }
  }
}
```


# Metaweb clustering
In this part, we apply SBM to infer trophic groups from metaweb

## Stochastic Block Model
```{r}
if(sbm_done==0){
  metaweb=alltaxa_analysis$Metaweb
  metaweb$print()
  ### We remove resource nodes ###
  metaweb$graph=delete_vertices(metaweb$graph,c("R1","R2","R3"))
  metaweb$print()
  ### Actual clustering ###
  sbm_out=troph_sbm(metaweb,file_raw_adjacency,file_raw_weighted_adjacency,file_sbm_groups,file_sbm_adjacency,minim=10)
  #### Get learnt parameters ###
  trophic_groups=sbm_out$tg
  sbm_adj=sbm_out$adj
  sbm_done=1
}else{
  trophic_groups=read.csv2(file_sbm_groups,row.names=1)
  sbm_adj=read.csv2(file_sbm_adjacency,row.names = 1)
}
```

## Functional diversity
Analyzing the functional composition of our communities
```{r}
trophic_groups=read.csv2(file_sbm_groups,row.names = 1)
func_occur=merge(agg_occur,trophic_groups,by.x="key",by.y="keys")
agg_func_occur=aggregate(func_occur[,pt_names]>0,by = list(trophGroup=func_occur$trophic_group),FUN = sum)
fac_func_occur=data.frame(t(agg_func_occur[,2:dim(agg_func_occur)[2]]))
colnames(fac_func_occur)<-agg_func_occur$trophGroup

rf=cbind(fac,redund_fct(agg_func_occur,pt_names,eta=c(0,1,2)))
rf_counts=cbind(fac,fac_func_occur)

rm(fac_func_occur)
rm(fac_func_occur)
rm(agg_func_occur)
rm(func_occur)

write.csv2(rf,file=file_redund_fct,row.names = F)
write.csv2(rf_counts,file_agg_func_occur,row.names = F)
```

## Trophic groups betadiversity
At point level (64 points) of SBM groups
```{r message=FALSE}
eta=c(0,1,2)
divs=c('P','L')
if(sbmbetapt_done==0){
  gList=pointlevel_analysis$SubNetworks
  nodenames=V(getMetaweb(gList))$name
  nodegroups=unlist(lapply(nodenames,
                           function(x) as.character(
                             subset(trophic_groups,keys==x)[1,"trophic_group"])))
  names(nodegroups)=nodenames
  
  for (div in c('P','L')){
    for (e in eta){
      sbmbetadiv_ptdf=betanet(gList = gList,
                              groups=nodegroups,
                              pt_names = lnames,
                              div=div,
                              eta=e,
                              file_beta = paste(file_sbmbeta_pt,"_",e,"_",div,".csv",sep=""))
    }
  }

  sbmbetapt_done=1
}else{
  sbmbetadiv_ptdf=read.csv2(file_sbmbeta_pt,row.names = 1)
}
x=pcoa_plot(D=as.dist(sbmbetadiv_ptdf),Y=NULL,corr="none")
```

<!-- Metaweb from SBM groups and projection to points -->
<!-- ```{r} -->
<!-- func_occur_pt<-aggregate_occur(agg_func_occur,pt_names,lgroups) -->
<!-- ### Get SBM adjacency matrix, create SBM metaweb, project on local plots/points/repets given agg_func_occur, compute beta_diversity ### -->
<!-- sbm_adj=read.csv2(file_sbm_adjacency,row.names = 1) -->
<!-- sbm_adj[sbm_adj<1E-3]<-0 -->
<!-- #sbm_adj[sbm_adj>=1E-3]<-1 -->
<!-- rownames(sbm_adj)<-colnames(sbm_adj) -->
<!-- mwsbm=graph_from_adjacency_matrix(as.matrix(sbm_adj),mode="directed",weighted=TRUE) -->
<!-- E(mwsbm)$cooccur<-E(mwsbm)$weight -->
<!-- E(mwsbm)$type<-rep("troph",vcount(mwsbm)) -->
<!-- plot(mwsbm) -->

<!-- sbm_mw=get.data.frame(mwsbm,what=c("both")) -->
<!-- sbm_analysis=meta_bionet_workflow(sbm_mw$edges,agg_func_occur,sbm_mw$vertices, -->
<!--                                name="BISE_SBM",pt_names, -->
<!--                                edge_ends=c("from","to"),edge_attributes=c("type","cooccur"), -->
<!--                                weight_att=2,type_att=1,min_prob=0,min_occur=1, -->
<!--                                verb=0) -->
<!-- ``` -->



<!-- ## Analyzing the trophic web of groups inferred by SBM -->
<!-- ```{r} -->
<!-- #### Add cluster assignments to metaweb nodes #### -->
<!-- cat("Adding cluster information to metaweb \n") -->
<!-- metaweb$graph=set_vertex_attr(graph=metaweb$graph,name="cluster", -->
<!--                               index = V(metaweb$graph), -->
<!--                               value=trophic_groups$trophic_group) -->

<!-- cat("Creating clustered metaweb \n") -->
<!-- trophic_groups$broad=taxa_metadata$broad -->
<!-- trophic_groups=merge(trophic_groups,taxa_metadata,by.x = 'keys',by.y = 'key') -->

<!-- cat("Creating trophic group nodes \n") -->
<!-- gt_nodes=lapply(1:argmax,function(x){ -->
<!--   metadata=subset(trophic_groups,trophic_group==x) -->
<!--   microhab=sapply(metadata[,c("surf","subsurf","soil")],max) -->
<!--   rank=as.character(unique(metadata[,"rank"])) -->
<!--   broad=as.character(unique(metadata[,"broad"])) -->
<!--   node=Taxon$new(gbif_id=x, -->
<!--                  label = paste('TG_',x,sep=""), -->
<!--                  broad=as.vector(broad), -->
<!--                  rank=as.vector(rank), -->
<!--                  microhab=microhab) -->
<!-- }) -->

<!-- cat("Creating edge list from SBM probabilities \n") -->
<!-- epsilon=1E-4 -->
<!-- sbm_el=melt(sbm_adj,varnames = c("resource","consumer"),value.name = "cooccur") -->
<!-- sbm_el$type="troph" -->
<!-- ### Discrétiser -->
<!-- sbm_el=sbm_el[!(sbm_el$cooccur<epsilon),] -->
<!-- group_metaweb=MetaWeb$new(mwebname="BISE_GT", taxa=gt_nodes,  -->
<!--                           edge_list=sbm_el[,c("resource","consumer")],  -->
<!--                           edge_metadata=sbm_el[,c("cooccur","type")]) -->

<!-- cat("Aggregating occurrences to groups in observation points \n") -->
<!-- #agg_occur_groups=merge(agg_occur,trophic_groups[,c("trophic_group","keys")],by.x="key",by.y="keys") -->
<!-- agg_occur_groups=agg_occur[,setdiff(colnames(agg_occur),"broad")] -->
<!-- agg_occur_groups$TG=unlist(lapply(agg_occur$key,function(x) subset(trophic_groups,keys==as.character(x))[1,"trophic_group"])) -->

<!-- observations=setdiff(colnames(agg_occur_groups),c("key","verbatimScientificName","TG")) -->
<!-- sum_occur_groups=aggregate(agg_occur_groups[,observations],by = list(TG=agg_occur_groups$TG),FUN = sum) -->

<!-- cat("Projecting group metaweb to observations \n") -->
<!-- gmin_occur=1 -->
<!-- group_subnets=lapply(observations, function(x){ -->
<!--   community_compos=as.character(sum_occur_groups[which(sum_occur_groups[,x]>=gmin_occur),"TG"]) -->
<!--   subnetwork=group_metaweb$project_metanetwork(sublist_taxa_keys = community_compos,net_name = x)  -->
<!--   return(subnetwork) -->
<!-- }) -->

<!-- cat("Selecting trophic groups containing light \n") -->

<!-- cat("Selecting trophic groups containing som \n") -->

<!-- cat("Selecting trophic groups containing bacteria decomposers \n") -->

<!-- cat("Selecting trophic groups containing fungi decomposers \n") -->

<!-- cat("Computing and saving energy pathways metrics for trophic groups on observation points \n") -->
<!-- group_topoenerg_metrics=lapply(group_subnets,topoenergetic_metrics) -->
<!-- cpt=1 -->
<!-- for (net in group_subnets){ -->
<!--   print(paste("cpt=",cpt)) -->
<!--   mets=topoenergetic_metrics(net) -->
<!--   cpt=cpt+1 -->
<!-- } -->
<!-- group_topometrics_df=do.call(rbind,lapply(group_topoenerg_metrics, as.data.frame)) -->
<!-- group_topometrics_df$observation_id=observations -->
<!-- ### Saving ### -->
<!-- write.csv2(topometrics_df,file="../../Applications/BISE/trophic_group_topo_energetics_metrics.csv",row.names = F) -->

<!-- ``` -->


